The Hot Hand Fallacy in Call of Duty Search and Destroy

### 2022-08-29 paper/paper.Rmd
everything but table 7 analysis

---
title: "The Hot Hand Fallacy in Call of Duty Search and Destroy"
authors:
  - name: Person One
    email: personone@domain.com
  - name: Person Two
    email: persontwo@domain.com
abstract: |
  Our research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, we find evidence supporting the hypothesis that round win probability can be modeled as a constant across rounds in the series, although not at the naive 50%. Second, we examine the proportion of round wins immediately following a win streak. Given streak length, series length, and series winner, we find no evidence that post-streak win proportion is significantly different than expected, suggesting that the perception of momentum at the team level is deceiving. Wald-Wolfowitz run tests also fail to provide evidence for the hot hand phenomenon.
bibliography: references.bib
keywords:
  - esports
  - Call of Duty
  - hot hand
  - runs test
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

[^5]: CoD has roughly gone through three eras of professional gaming: (1) Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World League (CWL), initiated in 2016; and (3) the 12-franchise CoD League (CDL), operating since 2020. The CDL has completed three year-long "seasons" as of August 2022.

    CoD is fairly unique compared to other esports in that it runs on an annual lifecycle, with releases coming in the late fall. A new game under the same brand---Call of Duty---is published every year by a rotating set of developers. Each new game bears resemblance to past ones, often introducing relatively small variations ("improvements") to graphics, game modes, and other facets of gameplay. During the CDL era, the games released have been Modern Warfare (2020), Cold War (2021) and Vanguard (2022).

@@

Sweeps (6-0 series) make up 4.7% of all series, while 6-5 series make up 23.7% of series. The observed offensive round win percentage across all rounds is $\bar{\tau}$ = 47.8%.[^8] Table \ref{tbl:o-win-prop-by-series-state} shows offensive round win percentages by series "state", i.e. the number of round wins by each team prior to an upcoming round. Offensive round win percentage is not quite constant, although never veers more than 10% from $\bar{\tau}$.

[^8]: Offensive round win percentage has been nearly constant across the three games during the CDL era: 1. 47.2% in MW (2020) 2. 47.9% in Cold War (2021) 3. 48.1% in Vanguard (2022)

@@

\end{table}
```

The professional competition format balances the frequency with which teams play offense or defense to start a series, to the extent possible. Thus, win percentages in certain states are subjected to minimal selection bias, i.e. bias due to better teams playing defense more frequently.[^9]

[^9]: At tournaments, higher-seeded ("better") teams choose whether they want to start on offense in the SnD series, played as the second or fifth game mode of the best-of-five matchup. Their opponent is assigned to start on offense in the SnD slot not chosen. Anecdotally, the better team tends to choose to play defense in the first round of the SnD series played as the second game mode in the best-of-five matchup, although this choice is not consistent across teams, or even with the same team over time.

@@

The concept of momentum goes hand-in-hand with a discussion of the nature of series outcomes.[^10] Two opposing fallacies are observed in the context of momentum: the "gambler's fallacy" (negative recency) and the "hot hand fallacy" (positive recency). Per Ayton et al. [-@ayton2004], negative recency is "the belief that, for random events, runs of a particular outcome ... will be balanced by a tendency for the opposite outcome", while positive recency is the expectation of observing future results that match recent results.

[^10]: We often use "streaks" and momentum interchangeably, but as Steeger et al. [-@steeger2021] note, momentum implies dependence between events, whereas streaking does not.

@@

1.  Both have economic aspects that can create clear advantages for one team in a round, given how prior rounds played out. CoD has no such equivalent, except for perhaps "score streaks", which infrequently occur.
2.  Teams play either offense or defense for many consecutive rounds. (The former has a 15-15-1 format and the latter uses a 12-12-1 format.) On the other hand, teams in CoD SnD rotate roles every round, analogous to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series for professional sports like the MLB, NBA, and NHL.[^11] While theoretically one might be able to account for inter-round win percentage relationships in any kind of format, such as a 5-5-1, the rotation of team sides every round is convenient for convincing ourselves that rounds could reasonably be modeled as i.i.d. Bernoulli trials.
3.  Both Counter-Strike and Valorant have overtime rules---a team must win by two rounds---which can make end-of-series sequences difficult to model properly. CoD SnD does not have overtime rules.

[^11]: 1-1-1-1-1-1-1 is not used today in these leagues, but it was at least once in each.

To our knowledge, there is no existing public statistical research on the CoD SnD format, beyond descriptive analysis on social media. While intra-round trends may be more directly applicable to teams looking for an advantage on their competition, broader investigation of a concept like the hot-hand fallacy in a sport where it has not yet been invetigated---particularly one that is less subjected to factors that may be difficult to control for, e.g. weather---should be useful as an inspiration for future researchers.

@@

In a best-of-$s$ format, assuming a constant round win probability $\phi$ (where $0 \leq \phi \leq 1$)[^12] for one team[^13], the expected proportion of series ending in $r$ rounds ($r \leq s$) is given by Equation \ref{eq:series-length}.

[^12]: If $\phi > 0.5$, we can say that this is the stronger team.

[^13]: $\phi$ and other symbols are selected so as to reserve symbols like $p$ for usage in other contexts without causing confusion for the reader. Upper-case symbol, e.g. $\Phi$, are consistently used in this paper to represent proportions, while lower-case symbols corresponding to upper-case symbols, e.g. $\phi$, are used to represent Bernoulli trial success rates. "Hats", e.g. $\hat{\Phi}$, are used to convey expectations, while bare symbols symbols convey observational data.

@@

```{=tex}
\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum_{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}
\end{equation}
```
in which $\hat{\Phi}(r)$ and $\Phi(r)$ represent the expected and observed round win proportions respectively, and $R = [6, 7, 8, 9, 10, 11]$ for CoD SnD. The null hypothesis is that the test statistic is chi-squared distributed, where the number of degrees of freedom is $\vert R \vert -1 = 5$ for Cod SnD. TODO: Make a note about $\alpha$

@@

Let us now consider round win proportions immediately following a streak of $k$ wins, given that the series lasts $r$ rounds. Removing our knowledge of a streak, we might model the Bernoulli round win probability as

```{=tex}
\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 := \begin{cases} 
  \frac{m}{r} & \text{team wins series}, \\ 
  \frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}
```
for $r \in R$ and $m$ from Equation \ref{eq:m}.

As implied by MS's Theorem 1, we should expect the proportion of round wins immediately following a streak of $k$ rounds wins for a series lasting $r$ rounds, $\hat{P}^{+|k,r}_{MS}$, to be strictly less than $p^{+|r}$. However, we should not mirror the equal proportions hypothesis testing performed by MS.

Our context is fundamentally different from that of MS, who focus on longitudinal data in controlled settings, unaffected by opposition.[^14] The number of trials is fixed in their experimental designs, but in CoD SnD, the number of rounds played is determined as a function of the max number of possible rounds ($s$) and whether or not the team wins the series. The Bernoulli trial success probability, i.e. single round win percentage in CoD SnD, is not independent of the opponent. Consequently, a statistical test of the difference in $\hat{P}^{+|k,r}$ and $\hat{P}^{-|kr}$, as performed by MS to evaluate their hypothesis regarding post-streak success rate, is not completely appropriate, although useful as a reference.

[^14]: GVT also perform statistical tests on shots from players in live games, i.e. "observational" data, but they note that their findings are likely affected by player shot selection in the face of tactics by the opposing team.

@@

One can consider another form of the expected proportion of rounds won immediately after a streak of $k$ round wins in a best-of-$s$ series given the final length of the series ($r$ rounds), the "notional" proportion $\hat{P}^{+|k,r}_0$. The Bernoulli round win probability underlying $\hat{P}^{+|k,r}_0$ is

```{=tex}
\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 := \begin{cases}
  \frac{m - k}{r - k} & \text{team wins series}, r > m; \\
  1 & \text{team wins series}, r = m; \\
  \frac{s - m - k}{r - k} & \text{team loses series}, r > m + k; \\
  0 & \text{team loses series}, r = m + k,
\end{cases}
\end{equation}
```

where $r \in R$ for $m$ from Equation \ref{eq:m}.

We can perform a two-tailed binomial test to evaluate the null hypothesis $\omega = \omega_0$ for the observed probability of success $\omega$ and a user-specified $\omega_0$ (where $0 \leq \omega_0 \leq 1$). If there are $r^+$ observed successes in a sample of $r$ trials, the probability of arriving at this expected number of successes is

```{=tex}
\begin{equation}\label{eq:binom}
\binom {r}{r^+} \omega^{r^+}(1-\omega)^{r-r^+}.
\end{equation}
```

We should expect that there should be $r * \omega_0$ successes.

TODO: Simplify this first sentence.

Treating the notional proportion $\hat{P}^{+|k,r}_0$ as the null $\omega_0$ and plugging in the observed proportion $P^{+|k,r}$ for $\omega$ in Equation \ref{eq:binom} (treating proportions as probabilities), we can evaluate the null hypothesis that the observed proportion of post-streak round wins is equal to the notional proportion, given streak length $k$ and series length $r$. If we can reject this null hypothesis, then we can consider team momentum, represented by post-streak success, plausible.

As a reference, we can perform the same binomial test for the MS's streak-selection-adjusted proportion, $\hat{P}^{+|k,r}_{MS}$. However, given the caveats mentioned before regarding applying MS's theorem to our setting, the results of such binomial tests should be heeded with caution.

@@

```{=tex}
\begin{equation}\label{eq:pwkri}
p_0(\text{win} | k, r, i) = p^{+|k,r,i}_0 = \begin{cases}
  \frac{m - k}{r - k} & \text{team wins series}, i \neq r; \\
  1 & \text{team wins series}, i = r; \\
  \frac{s - m - k}{r - k} & \text{team loses series}, i \neq r, r > m + k; \\
  0 & \text{team loses series}, i = r \text{ or } r = m + k.
\end{cases}
\end{equation}
```
Again, we can apply a binomial test to evaluate the hypothesis that the expected proportion--$\hat{P}^{+|k,r,i}_0$ and $\hat{P}^{+|k,r,i}_{MS}$ separately--is equal to the observed proportion $P^{+|k,r,i}$.

@@

Stepping back from the perspective of a single team's round win probability when streaking, one can attempt to detect the hot hand phenomenon more broadly with a Wald-Wolfowitz runs test. Under the null hypothesis, the number of runs in a sequence of $r$ trials is a random variable that can take on values $+$ or $-$ and arrive at $r^+$ successes ($r^- = r - r^+$ failures), with the following mean $\mu$ and variance $\sigma^2$:

```{=tex}
\begin{equation}\label{eq:ww}
\mu = \frac{2r^{+}r^{-}}{r} + 1, \sigma^2 = \frac{(\mu-1)(\mu-2)}{r-1}.
\end{equation}
```
To incorporate our findings regarding constant round win probability, we can specify that $\Pr(r^+) = \phi$.

One can subset the observed series sequences, $\zeta(r)$, to those that violate the null hypothesis for the runs test and perform a test of equal proportions, where the null is that the observed proportion of a sequence relative to all possible sequences, $P^\zeta$, is equal to the expected proportion of the sequence relative to all possible sequences, $\hat{P}^\zeta$. The test statistic is

```{=tex}
\begin{equation}\label{eq:prop}
z = \frac{P^\zeta - \hat{P}^\zeta}{\sqrt{P^\zeta_{\delta} (1 - P^\zeta_{\delta}) (1 / N^\zeta + 1 / \hat{N}^\zeta)  } }
\end{equation}
```
where

$$
P^\zeta_{\delta} = \frac{P^\zeta - \hat{P}^\zeta}{N^\zeta - \hat{N}^\zeta}
$$

and where $N^\zeta$ and $\hat{N}^\zeta$ are the observed and expected number of sequences respectively. The null hypothesis is that $z$ is normally distributed. If we can reject the null hypothesis for such sequences and build an argument in support of the hot hand effect.

@@

Using Equation \ref{eq:chi-squ}, we find that $\chi^2 = 16.0$ ($p$-value of <0.01) for $\phi_0 = 0.5$. Thus, we can comfortably reject the constant probability hypothesis for the null $\phi_0 = 0.5$ at a confidence level of $\alpha = 0.05$.

@@

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion of series lasting $r$ rounds given $\phi_0 = 0.5$, $\hat{\Phi}_0(r)$, as well as the observed proportions, $\Phi(r)$.

```{=tex}
\begin{table}
\caption{The expected proportion of CoD SnD series lasting $r$ rounds, $\hat{\Phi}_0(r)$, under the assumption that each team has a constant round win probability $\phi_0 = 0.5$. Additionally, the observed frequencies for CoD SnD shown as a count $N(r)$ and as a proportion $\Phi(r)$ of all series ($\sum_{r \in R} N(r)$, where $r \in R = [6, 7, 8, 9, 10, 11]$).}

@@

\end{table}
```
Table \ref{tbl:mosteller-methods-results} shows the alternate values for the constant round win probability assumptions, each approximately equal to 0.575. When applying Equation \ref{eq:chi-squ}, each results in a $\chi^2$ value for which we cannot reject the constant probability null hypothesis.

@@

\caption{Alternate estimates of the constant probability ($\phi$) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive ($\phi_0 = 0.5$).}

@@

0. Naive & 0.5000 & 16.0 ($\leq$ 0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum ($\chi^2$) & 0.5775 & 3.5 (0.62) \\

@@

\end{table}
```
Table \ref{tbl:alternative-constant-ps} shows the new $\hat{\Phi}(r)$ when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $\phi_2 = 0.575$, resulting in a new set of expected proportions of series lasting $r$ rounds $\hat{\Phi}_2(r)$.[^15] We observe that $\hat{\Phi}_2(r)$ more closely matches $\Phi(r)$ than $\hat{\Phi}_0(r)$.

[^15]: The method of moments and minimum $\chi^2$ estimates for $\phi$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $\phi$ (since they are all $\approx 0.575$).

```{=tex}
\begin{table}
\caption{The observed proportion of CoD SnD series, $\Phi(r)$, ending in $r$ rounds, compared to the expected proportion, $\hat{\Phi}_0(r)$, under the naive assumption $\phi_0 = 0.5$ and the expected proportion, $\hat{\Phi}_2(r)$, under the maximum likelihood estimate, $\phi_2 = 0.575$, for constant round win probability.}

\centering
\begin{tabular}{rrrrrr}
\toprule
$r$ & $\Phi(r)$ & $\hat{\Phi}_0(r)$ = 0.5 & $\hat{\Phi}_2(r)$ = 0.575 \\
\midrule

@@

\end{table}
```
Observing the null hypothesis acceptance shown in Table \ref{tbl:mosteller-methods-results}, along with the observation that $\hat{\Phi}_2(r)$ reasonably matches $\Phi(r)$, we can say that the constant round win probability assumption is valid in CoD SnD series with the appropriate choice of $\phi$ ($\approx 0.575$).

@@

  cod_round_win_prop_after_z_with_probs |> 
    mutate(
      notional_p_value = vectorized_binom_test(n, total, p = notional_prop)$p.value,
      ms_p_value = vectorized_binom_test(n, total, p = ms_prop)$p.value
    )
}

@@

cod_round_win_prop_after_k_wins <- tibble(win_streak = 2L:5L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)
cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value < 0.05)
```

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], we focus on streaks of three round wins. Table \ref{tbl:pwkr} compares the notional and MS expected proportions, $\hat{P}^{+|k,r}_0$ and $\hat{P}^{+|k,r}_{MS}$ respectively, with the observed round win proportion, $P^{+|k,r}$, following streaks of $k=3$ round wins given the series outcome.

@@

\caption{Given the round win streak $k=3$, the length of the series ($r$ rounds), and the series winner, the observed count of rounds wins, $r^{+|k=3,r}$, and proportion of round wins, $P^{+|k=3,r}$, among $N^{k=3,r}$ instances where a team could win after the streak ($P^{+|k=3,r} = \frac{r^{+|k=3,r}}{N^{k=3,r}}$). Additionally, the notional and MS expected proportions, $\hat{P}^{+|k=3,r}_0$ and $\hat{P}^{+|k=3,r}_{MS}$ respectively.}

@@

cod_round_win_prop_after_k_wins_i_rounds <- init_cod_round_win_prop_after_k_wins_i_rounds |>
  group_by(win_streak, win_round, n_rounds, round) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  ungroup() |> 
  mutate(
    notional_p_value = vectorized_binom_test(n, total, p = notional_prop)$p.value |> round(2)
  )

>>>>

---
title: "The Hot Hand Fallacy in Call of Duty Search and Destroy"
link-citations: true
authors:
  - name: Person One
    email: personone@domain.com
  - name: Person Two
    email: persontwo@domain.com
abstract: |
  Our research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). First, we find evidence supporting the hypothesis that round win probability can be modeled as a constant across rounds in the series, although not at the naive 50%. Second, we examine the proportion of round wins immediately following a win streak. Given streak length, series length, and series winner, we find no evidence that post-streak win proportion is significantly different than expected, suggesting that the perception of momentum at the team level is deceiving. Wald-Wolfowitz run tests also fail to provide evidence for the hot hand phenomenon.
bibliography: references.bib
keywords:
  - esports
  - Call of Duty
  - hot hand
  - runs test
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

[^5]: Since the release of "Call of Duty: Modern Warfare" in fall 2019, professional CoD has been orchestrated by the 12-franchise CoD League (CDL). The CDL has completed three year-long "seasons" as of August 2022.

@@

CoD is fairly unique compared to other esports in that it runs on an annual lifecycle, with releases coming in the late fall. A new game under the same brand---"Call of Duty"---is published every year by a rotating set of developers. Each new game bears resemblance to past ones, often introducing relatively small variations ("improvements") to graphics, game modes, and other facets of gameplay. During the CDL era, the games released have been "Modern Warfare" (2020 season), "Cold War" (2021 season) and "Vanguard" (2022 season).

Sweeps (6-0 series) make up 4.7% of all series, while 6-5 series make up 23.7% of series. The observed offensive round win percentage[^8] across all rounds is $\bar{\tau}$ = 47.8%.[^9] Table \ref{tbl:o-win-prop-by-series-state} shows offensive round win percentages by series "state", i.e. the number of round wins by each team prior to an upcoming round. Offensive round win percentage is not quite constant, although never veers more than 10% from $\bar{\tau}$.

[^8]: We use the terms "percentage", "probability", and "proportion" synonymously throughout this paper when discussing round win rates. For the most part, we restrict "percentage" usage to observed values and "probability" usage to expected values, and we use "proportion" to refer to either observed or expected values.

[^9]: Offensive round win percentage has been nearly constant across the three games during the CDL era: 47.2% in Modern Warfare; 47.9% in Cold War; and 48.1% in Vanguard

@@

\end{table}
```
The professional competition format balances the frequency with which teams play offense or defense to start a series, to the extent possible. Thus, win percentages in certain states are subjected to minimal selection bias, i.e. bias due to better teams playing defense more frequently.[^10]

[^10]: At tournaments, higher-seeded ("better") teams choose whether they want to start on offense in the SnD series, played as the second or fifth game mode of the best-of-five matchup. Their opponent is assigned to start on offense in the SnD slot not chosen. Anecdotally, the better team tends to choose to play defense in the first round of the SnD series played as the second game mode in the best-of-five matchup, although this choice is not consistent across teams, or even with the same team over time.

@@

The concept of momentum goes hand-in-hand with a discussion of the nature of series outcomes.[^11] Two opposing fallacies are observed in the context of momentum: the "gambler's fallacy" (negative recency) and the "hot hand fallacy" (positive recency). Per Ayton et al. [-@ayton2004], negative recency is "the belief that, for random events, runs of a particular outcome ... will be balanced by a tendency for the opposite outcome", while positive recency is the expectation of observing future results that match recent results.

[^11]: We often use "streaks" and momentum interchangeably, but as Steeger et al. [-@steeger2021] note, momentum implies dependence between events, whereas streaking does not.

@@

1.  Both have economic aspects that can create clear advantages for one team in a round, given how prior rounds played out. CoD has no such equivalent, except for perhaps "score streaks", which infrequently occur.
2.  Teams play either offense or defense for many consecutive rounds. (Counter-Strike has a 15-15-1 format and Valorant has a 12-12-1 format.) On the other hand, teams in CoD SnD rotate attacking and defending roles every round, minimizing the effect of map asymmetry on streaks of wins.[^12]
3.  Both Counter-Strike and Valorant have overtime rules---a team must win by two rounds---which can make end-of-series sequences difficult to model properly. CoD SnD does not have overtime rules.

[^12]: Professional CoD has a set of maps (at least three, depending on the season) on which teams choose to play SnD. Most are asymmetric, introducing slight advantages for the offense or defense that do not exist on other maps. Counter-Strike and Valorant also have their own sets of maps for their analogous SnD modes.

To our knowledge, there is no existing public statistical research on the CoD SnD format, beyond descriptive analysis on social media. While intra-round trends may be more directly applicable to teams looking for an advantage on their competition, broader investigation of a concept like the hot-hand fallacy in a sport where it has not yet been investigated---particularly one that is less subjected to factors that may be difficult to control for, e.g. weather---should be useful as an inspiration for future researchers.

@@

In a best-of-$s$ format, assuming a constant round win probability $\phi$ (where $0 \leq \phi \leq 1$) for a given team, the expected proportion of series ending in $r$ rounds (where $r \leq s$) is given by Equation \ref{eq:series-length}.

@@

```{=tex}
\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum_{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}
\end{equation}
```
where $\hat{\Phi}(r)$ and $\Phi(r)$ represent the expected and observed round win proportions respectively, and where $R = [6, 7, 8, 9, 10, 11]$ for CoD SnD. The null hypothesis is that the test statistic is chi-squared distributed, where the number of degrees of freedom is $\vert R \vert -1 = 5$ for Cod SnD. We use a significance level of $\alpha = 0.05$ for this test and all others in this paper.

@@

Let us now consider round win proportions immediately following a streak of $k$ wins, given that the series lasts $r$ rounds. Removing our knowledge of a streak, we might expect the win proportion for a round to be

```{=tex}
\begin{equation}\label{eq:pwr}
\hat{P}_0(\text{win} | r) = \hat{P}^{+|r}_0 := \begin{cases} 
\frac{m}{r} & \text{team wins series}, \\ 
\frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}
```
for $r \in R$ and $m$ from Equation \ref{eq:m}.

As implied by MS's Theorem 1, we should expect the proportion of round wins immediately following a streak of $k$ rounds wins for a series ending in $r$ rounds, $\hat{P}^{+|k,r}_{MS}$, to be strictly less than the observed proportion of round wins given the series length, but not given knowledge of the potential existence of a streak, $P^{+|r}$, i.e. the analogue of the expected proportion in Equation \ref{eq:pwr} for the observed data. However, our context is fundamentally different from that of MS, who evaluate data in controlled settings, unaffected by opposition.[^13]

[^13]: GVT also perform statistical tests on shots from players in live games, i.e. "observational" data, but they note that their findings are likely affected by player shot selection influenced by the opposing team's defensive strategy. This is similar to our setting.

The number of trials is fixed in their experimental designs; in CoD SnD, however, the number of rounds played is determined as a function of the max number of possible rounds ($s$) and whether or not the team wins the series. The single round win percentage in CoD SnD---analogous to a player's shooting percentage in MS's analysis of basketball players---is not independent of the opponent. Consequently, the i.i.d. assumption of MS's theorem regarding post-streak success may or may not be valid. Nonetheless, we still leverage $\hat{P}^{+|k,r}_{MS}$ as a reference for our results.

@@

Let us consider another form of the expected proportion of rounds won immediately after a streak of $k$ round wins in a best-of-$s$ series ending in $r$ rounds: the "notional" proportion

```{=tex}
\begin{equation}\label{eq:pwkr}
\hat{P}_0(\text{win} | k, r) = \hat{P}^{+|k,r}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, r > m; \\
1 & \text{team wins series}, r = m; \\
\frac{s - m - k}{r - k} & \text{team loses series}, r > m + k; \\
0 & \text{team loses series}, r = m + k,
\end{cases}
\end{equation}
```
where $r \in R$ and $m$ from Equation \ref{eq:m}.

We can perform a two-tailed binomial test to evaluate the null hypothesis $\omega = \omega_0$ for the observed proportion of successes, $\omega$, and a null proportion, $\omega_0$ (where $0 \leq \omega_0 \leq 1$). The probability of exactly $r^+$ successes (and $r^- = r - r^+$ failures) in a sample of $r$ trials is

```{=tex}
\begin{equation}\label{eq:binom}
\binom {r}{r^+} \omega^{r^+}(1-\omega)^{r^-},
\end{equation}
```
where $\omega = \frac{r^+}{r}$ is the observed proportion of successes. If we can reject the null hypothesis---treating $P^{+|k,r}$ and $\hat{P}^{+|k,r}_0$ as $\omega$ and $\omega_0$ respectively---then we might consider team momentum plausible. Since we perform this test for many combinations of $k$ and $r$, we adjust the $p$-values with the Benjamini and Yekutieli (BY) false discovery procedure [-@benjamini2001].

As a reference, we can perform the same binomial test for MS's proportion, $\hat{P}^{+|k,r}_{MS}$, which adjusts for streak selection bias. However, given the caveats mentioned before regarding the application of MS's theorem to our setting, the results of such binomial tests should be heeded with caution.

@@

```{=tex}
\begin{equation}\label{eq:pwkri}
\hat{P}_0(\text{win} | k, r, i) = \hat{P}^{+|k,r,i}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, i \neq r; \\
1 & \text{team wins series}, i = r; \\
\frac{s - m - k}{r - k} & \text{team loses series}, i \neq r, r > m + k; \\
0 & \text{team loses series}, i = r.
\end{cases}
\end{equation}
```
Again, we can apply a binomial test to evaluate the hypothesis that the expected proportion, $\hat{P}^{+|k,r,i}_0$, is equal to the observed proportion, $P^{+|k,r,i}$.

@@

We can attempt to detect the hot hand phenomenon more broadly by focusing on non-random sequences identified by the Wald-Wolfowitz runs test. Note that the probability of success and failure need not be equal; thus, we can incorporate our findings regarding constant round win probability, specifying that the probability of success is $\phi_2 = 0.575$.

Under the null hypothesis for the Wald-Wolfowitz runs test, the number of runs in a sequence of $r$ trials is a random variable that can take on values $+$ or $-$ and arrive at $r^+$ successes (and $r^-$ failures), with the following mean $\mu_r$ [^14] and variance $\sigma_r^2$:

[^14]: We calculate $\mu$ and $\sigma^2$ for each number of rounds $r \in R$.

```{=tex}
\begin{equation}\label{eq:ww}
\mu_r = \frac{2r^{+}r^{-}}{r} + 1, \sigma_r^2 = \frac{(\mu_r-1)(\mu_r-2)}{r-1}.
\end{equation}
```
The null hypothesis is that the test statistic,

```{=tex}
\begin{equation}\label{eq:wwz}
z_r = \frac{r^* - \mu_r}{\sigma_r},
\end{equation}
```
where $r^*$ is the number of runs, is normally distributed.

We can subset the observed series sequences, $\zeta_j$, to those that violate the null hypothesis for the runs test and perform a test of equal proportions, where the null hypothesis is that the observed proportion of a sequence relative to all possible sequences, $P^{\zeta}_j$ ( $\frac{N^\zeta_j}{\sum_j N^\zeta }$), is equal to the expected proportion of the sequence relative to all possible sequences, $\hat{P}^\zeta_j$. The test statistic is

```{=tex}
\begin{equation}\label{eq:prop}
z^\zeta_j = \frac{P^\zeta_j - \hat{P}^\zeta_j}{\sqrt{P^\zeta_{j,\delta} (1 - P^\zeta_{j,\delta}) (1 / N^\zeta_j + 1 / \hat{N}^\zeta_j)  } }
\end{equation}
```
where $N^\zeta_j$ and $\hat{N}^\zeta_j$ respectively represent the observed and expected number of occurrences of sequence $\zeta_j$, and where $P^\zeta_{j,\delta} = (P^\zeta_j - \hat{P}^\zeta_j) / ({N^\zeta_j - \hat{N}^\zeta_j})$. The null hypothesis is that the test statistic is normally distributed. If we can reject the null hypothesis for such sequences, we can build an argument in support of the hot hand effect.[^15]

[^15]: As with the binomial tests for the notional and MS post-streak proportions, we adjust the $p$-values with the BY correction.

@@

Using Equation \ref{eq:chi-squ}, we find that $\chi^2 = 16.0$ ($p$-value of \<0.01) for $\phi_0 = 0.5$. Thus, we can comfortably reject the constant probability hypothesis for the naive $\phi_0 = 0.5$.

@@

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion of series ending in $r$ rounds given $\phi_0 = 0.5$, $\hat{\Phi}_0(r)$, as well as the observed proportions, $\Phi(r)$.

```{=tex}
\begin{table}
\caption{The observed frequencies of CoD SnD series ending in $r$ rounds shown as a count $N(r)$ and as a proportion $\Phi(r)$ of all series ($\sum_{r \in R} N(r)$, where $r \in R = [6, 7, 8, 9, 10, 11]$). Additionally, the expected proportion, $\hat{\Phi}_0(r)$, under the assumption that each team has a constant round win probability $\phi_0 = 0.5$.}

@@

\end{table}
```
Table \ref{tbl:mosteller-methods-results} shows the alternate values for the constant round win probability assumption, each approximately equal to 0.575. When applying Equation \ref{eq:chi-squ}, each results in a $\chi^2$ value for which we cannot reject the constant probability null hypothesis.

@@

\caption{Alternate estimates of the constant round win probability ($\phi$) and their corresponding $\chi^2$ values, applying the three methods suggested by Mosteller to CoD SnD. Additionally, the naive constant round win probability $\phi_0$ and its $\chi^2$.}

@@

$\phi_0$: Naive & 0.5000 & 16.0 ($\leq$ 0.01) \\
$\phi_1$: Method of moments & 0.5725 & 3.6 (0.6) \\
$\phi_2$: Maximum likelihood & 0.5750 & 3.5 (0.62) \\
$\phi_3$: Minimum $\chi^2$ & 0.5775 & 3.5 (0.62) \\

@@

\end{table}
```
Table \ref{tbl:alternative-constant-ps} shows the expected proportions of series ending in $r$ rounds, $\hat{\Phi}_2(r)$, when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $\phi_2 = 0.575$.[^16] We observe that $\hat{\Phi}_2(r)$ more closely matches $\Phi(r)$ than $\hat{\Phi}_0(r)$.

[^16]: The method of moments and minimum $\chi^2$ estimates for $\phi$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $\phi$ (since they are all $\approx 0.575$).

\caption{The observed proportion of CoD SnD series ending in $r$ rounds, $\Phi(r)$, compared to the expected proportion, $\hat{\Phi}_2(r)$, under the maximum likelihood estimate, $\phi_2 = 0.575$. Additionally, the expected proportion, $\hat{\Phi}_0(r)$, under the naive assumption $\phi_0 = 0.5$ is repeated from Table \ref{tbl:prob-series-lasts-r-rounds} for comparison.}

\centering
\begin{tabular}{rrrr}
\toprule
$r$ & $\Phi(r)$ & $\hat{\Phi}_0(r)$ = 0.5 & $\hat{\Phi}_2(r)$ = 0.575 \\
\midrule

@@

\end{table}
```
The observation that $\hat{\Phi}_2(r)$ reasonably matches $\Phi(r)$ supports the implication from the failure to reject the null hypothesis for $\phi_2$ (shown in Table \ref{tbl:mosteller-methods-results}): the constant round win probability assumption can be valid for CoD SnD with the appropriate choice of $\phi$ ($\approx 0.575$).

@@

  cod_round_win_prop_after_z_with_probs |> 
    mutate(
      notional_p_value = vectorized_binom_test(n, total, p = notional_prop)$p.value,
      ms_p_value = vectorized_binom_test(n, total, p = ms_prop)$p.value,
      notional_p_value_adj = notional_p_value |> p.adjust(method = 'BY')
    )
}

@@

init_cod_round_win_prop_after_k_wins <- tibble(win_streak = 2L:5L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)

cod_round_win_prop_after_k_wins <- init_cod_round_win_prop_after_k_wins |> 
  filter(win_round) |> 
  mutate(
    notional_p_value_adj = notional_p_value |> p.adjust(method = 'BY'),
    ms_p_value_adj = ms_p_value |> p.adjust(method = 'BY')
  )

cod_round_win_prop_after_k_wins |> filter(win_round, notional_p_value_adj < 0.05)
cod_round_win_prop_after_k_wins |> filter(win_round, ms_p_value_adj < 0.05)
```

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], we focus on streaks of three round wins. Table \ref{tbl:pwkr} compares the observed round win proportion, $P^{+|k=3,r}$, with the notional proportion, $\hat{P}^{+|k=3,r}$, following streaks of $k=3$ round wins. The MS proportion, $\hat{P}^{+|k=3,r}_{MS}$, is also shown as a reference.

@@

\caption{For streaks of $k=3$ round wins in CoD SnD series ending in $r$ rounds, the observed count of rounds wins, $r^{+|k=3,r}$, and proportion of round wins, $P^{+|k=3,r}$, among $N^{k=3,r}$ chances ($P^{+|k=3,r} = r^{+|k=3,r} / N^{k=3,r}$). Additionally, the notional and MS expected proportions, $\hat{P}^{+|k=3,r}_0$ and $\hat{P}^{+|k=3,r}_{MS}$, respectively.
}

@@

cod_round_win_prop_after_k_wins_i_rounds <- init_cod_round_win_prop_after_k_wins_i_rounds |>
  group_by(win_streak, win_round, n_rounds, round) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  ungroup() |> 
  mutate(
    notional_p_value = vectorized_binom_test(n, total, p = notional_prop)$p.value
  ) |> 
  filter(win_round) |> 
  mutate(
    notional_p_value_adj = notional_p_value |> p.adjust(method = 'BY')
  )
